Something about Performance, 2D Component of NEMO. (TBC)

### 2015-02-25 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
Tweak table containing perf. data for opt. stages and add text describing key features for each compiler.

\subsection{Addition of `safe-address' directives}

@@

\subsection{In-line Momentum kernel bodies into middle-layer code}

@@

\subsection{In-line the Continuity kernel}

@@

\subsection{In-line remaining kernels (BCs and time-update)}

@@

\subsection{In-line field-copy operations}

@@

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Performance (millions of points updated per second) for
  the $256^2$ case after each code transformation.}
\label{TABLE_opt_breakdown}
\centering
\begin{tabular}{l|c|c|c}
\hline
Compiler:           & Gnu & Intel & Cray  \\
\hline
CPU:                & \multicolumn{3}{c}{Ivy Bridge}  \\
\hline
Original         & 3.87 & 12.2 & 6.83 \\
Vanilla \psykal  & 3.07 & 6.55 & 6.73 \\
Safe-loop        & 3.07 & 6.55 & 6.95 \\
In-line Momentum & 4.31 & 10.7 & 6.93 \\
SIMD Momentum    & 4.31 & 10.7 & 11.8 \\
\parbox{2.5cm}{\raggedright Arrays from read-only objects} & 4.31 & 11.3 & 11.8 \\
In-line Continuity         & 4.83 & 11.8 & 11.6 \\
In-line remaining kernels  & 5.89 & 12.0 & 11.5 \\
In-line field copies       & 5.92 & 12.5 & 11.4 \\
\hline
\parbox{2.5cm}{\%-slow-down of best {\it c.f.} original} & -34.62 & -2.39 & -42.22 \\
\hline
\end{tabular}
\end{table}

\begin{figure}[!t]
\centering
\includegraphics[width=90mm]{opt_stages_256}
\caption{Performance of the \psykal version of nemolite2d for the
  $256^{2}$ domain at each stage of optimisation. Results are given as
  a percentage of the performance of the original code for each
  compiler/CPU combination. A figure greater than 100\% indicates that
  the \psykal version performs better than the original.}
\label{FIG_opt_stages_256}
\end{figure}

>>>>

\subsection{Addition of `safe-address' directives}
\label{sec_safe_address}

@@

\subsection{In-line Momentum kernel bodies into middle-layer code}
\label{sec_inline_mom}

@@

\subsection{In-line the Continuity kernel}
\label{sec_cont_inline}

@@

\subsection{In-line remaining kernels (BCs and time-update)}
\label{sec_inline_all_kernels}

@@

\subsection{In-line field-copy operations}
\label{sec_inline_fldcopy}

@@

Looking at the results for the Gnu compiler (and the Ivy Bridge CPU)
first, all of the steps-up in performance correspond to kernel
in-lining. None of the other transformations had any effect on the
performance of the compiled code. In fact, simply in-lining the two
kernels associated with the Momentum section was enough to improve
upon the performance of the original code.

With the Intel compiler, the single largest performance increase is
again due to kernel in-lining (of the Momentum kernels). This is
because the compiler does a much better job of SIMD vectorising the
loops involved than it does when it first has to in-line the kernel
itself (as evidenced by its own diagnostic output - see
Section~\ref{sec_inline_mom}). However, although this gives a
significant performance increase it is not sufficient to match the
performance of the original version. This is only achieved by
in-lining every kernel and making the lack of data dependencies
between arrays accessed from different field objects more explicit.

The Cray compiler is distinct from the other two in that kernel
in-lining does not give any performance benefit and in fact, for the
smaller kernels, it can actually hurt performance. Thus the key
transformation is to encourage the compiler to SIMD vectorise the
Momentum section via a compiler-specific directive (without this it
concludes that such vectorisation would be inefficient). Of the other
transformations, only the addition of the compiler-specific
safe\_address directive (see Section~\ref{sec_safe_address}) was found
to give any benefit.

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Performance (millions of points updated per second) for the
  $256^2$ case after each code transformation. Where an optimisation
  utilises compiler-specific directives then performance figures for
  the other compilers are omitted.}
\label{TABLE_opt_breakdown}
\centering
\begin{tabular}{l|c|c|c}
\hline
Compiler:           & Gnu & Intel & Cray  \\
\hline
CPU:                & \multicolumn{3}{c}{Ivy Bridge}  \\
\hline
Original         & 3.87 & 12.2 & 6.83 \\
Vanilla \psykal  & 3.07 & 6.55 & 6.73 \\
Safe-address     & --   & --   & 6.95 \\
In-line Momentum & 4.31 & 10.7 & 6.93 \\
SIMD Momentum    & --   & --   & 11.8 \\
\parbox{2.5cm}{\raggedright Grid data from read-only objects} & 4.31 & 11.3 & 11.8 \\
In-line Continuity         & 4.83 & 11.8 & 11.6 \\
In-line remaining kernels  & 5.89 & 12.0 & 11.5 \\
In-line field copies       & 5.92 & 12.5 & 11.4 \\
\hline
\parbox{2.5cm}{\raggedright \%-speed-up of best {\it c.f.} original} & 34.62 & 2.39 & 42.22 \\
\hline
\end{tabular}
\end{table}

\begin{figure}[!t]
\centering
\includegraphics[width=90mm]{opt_stages_256}
\caption{Performance of the \psykal version of nemolite2d for the
  $256^{2}$ domain at each stage of optimisation. The first (black)
  bar of each cluster gives the performance of the original version of
  nemolite2d for that compiler/CPU combination.}
\label{FIG_opt_stages_256}
\end{figure}
